---
title: "An R interface to the Enrichr database"
author: "Wajid Jawaid"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: enrichr.bib
vignette: >
  %\VignetteIndexEntry{An R interface to the Enrichr database}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Installation

**enrichR** can be installed from Github or from CRAN.

## Github

```{r, echo = FALSE, eval = TRUE}
websiteLive <- TRUE
```

```{r, echo = TRUE, eval = FALSE}
library(devtools)
install_github("wjawaid/enrichR")
```

## CRAN

The package can be downloaded from CRAN using:

```{r, echo = TRUE, eval = FALSE}
install.packages("enrichR")
```

# Usage example

**enrichR** provides an interface to the Enrichr database [@kuleshov_enrichr:_2016] hosted at https://maayanlab.cloud/Enrichr/.  

By default human genes are selected otherwise select your organism of choice. (This functionality was contributed by Alexander Blume)

## Initialising connection to Enrichr website

```{r, echo = TRUE, eval = TRUE}
library(enrichR)
websiteLive <- getOption("enrichR.live")
if (websiteLive) {
    listEnrichrSites()
    setEnrichrSite("Enrichr") # Human genes   
}
```

## Select gene-set libraries

List all available databases from Enrichr.

```{r, echo = TRUE, eval = TRUE}
if (websiteLive) dbs <- listEnrichrDbs()
```

```{r, echo = TRUE, eval = TRUE}
if (websiteLive) head(dbs)
```

Select the 2023 GO databases.

```{r, echo = TRUE, eval = TRUE}
dbs <- c("GO_Molecular_Function_2023", "GO_Cellular_Component_2023", 
	 "GO_Biological_Process_2023")
```

## Perform analysis

### Without background

Query with `enrichr` using genes available from the package.

```{r, echo = TRUE, eval = TRUE}
# Load example input genes
data(input)
length(input)
head(input)

if (websiteLive) {
    enriched <- enrichr(input, dbs)
}
```

Now view the `"GO_Biological_Process_2023"` results from `enriched`.

```{r, echo = TRUE, eval = FALSE}
if (websiteLive) head(enriched[["GO_Biological_Process_2023"]])
```

```{r, echo = FALSE, results = 'asis'}
success <- websiteLive & (length(enriched) >= 3)
success <- success & all(dim(enriched[["GO_Biological_Process_2023"]]) > 2)
if (success) {
    x <- head(enriched[["GO_Biological_Process_2023"]])
    x[,1] <- gsub(":", "&#58;", x[,1])
    knitr::kable(x)
}
```

### With background

You can now try adding a background to `enrichr`.

```{r, echo = TRUE, eval = TRUE}
# Load example background
data(background)
length(background)
head(background)

if (websiteLive) {
    enriched2 <- enrichr(input, dbs, background = background)
}
```

Now view the `"GO_Biological_Process_2023"` results from `enriched2`.

```{r, echo = TRUE, eval = FALSE}
if (websiteLive) head(enriched2[["GO_Biological_Process_2023"]])
```

```{r, echo = FALSE, results = 'asis'}
success <- websiteLive & (length(enriched2) >= 3)
success <- success & all(dim(enriched2[["GO_Biological_Process_2023"]]) > 2)
if (success) {
    x <- head(enriched2[["GO_Biological_Process_2023"]])
    x[,1] <- gsub(":", "&#58;", x[,1])
    knitr::kable(x)
}
```

By default, the results table from analysis with a background does not have the 'Overlap' column.
We can calculate the annotated genes in each term from GMT files and 
replace the 'Rank' column with 'Overlap' by setting `include_overlap = TRUE`.

```{r, echo = TRUE, eval = TRUE}
if (websiteLive) {
    # Replace 'Rank' with 'Overlap' column, calculated with GMT files
    enriched3 <- enrichr(input, dbs, background = background, include_overlap = TRUE)
}
```

Now view the `"GO_Biological_Process_2023"` results from `enriched3`.

```{r, echo = TRUE, eval = FALSE}
if (websiteLive) head(enriched3[["GO_Biological_Process_2023"]])
```

```{r, echo = FALSE, results = 'asis'}
success <- websiteLive & (length(enriched3) >= 3)
success <- success & all(dim(enriched3[["GO_Biological_Process_2023"]]) > 2)
if (success) {
    x <- head(enriched3[["GO_Biological_Process_2023"]])
    x[,1] <- gsub(":", "&#58;", x[,1])
    knitr::kable(x)
}
```

## Visualise results

Plot Enrichr `GO_Biological_Process_2023` output. (Plotting function contributed by I-Hsuan Lin)

```{r, echo = TRUE, eval = TRUE, fig.width = 8, fig.height = 5, fig.align = "center", dpi = 100}
if (websiteLive) {
    plotEnrich(enriched[["GO_Biological_Process_2023"]], showTerms = 20, numChar = 40, 
	       y = "Count", orderBy = "P.value")
}
```

## Export results

Export results to text files or Excel.

```{r, echo = TRUE, eval = FALSE}
# To text files
printEnrich(enriched)

# To Excel
printEnrich(enriched, outFile = "excel")

```

# References
